---
title: "Week 6 assignment"
author: "Harish"
date: "2022-10-10"
output: html_document
---

```{r, message=FALSE}
library(ggplot2) 
library(dplyr) 

train <- read.csv('train.csv')
test <- read.csv('test.csv')
# test doesn't have outcome - Survived

str(train)

```
Variable Name | Description
--------------|-------------
Survived      | Survived (1) or died (0)
Pclass        | Passenger's class
Name          | Passenger's name
Sex           | Passenger's sex
Age           | Passenger's age
SibSp         | Number of siblings/spouses aboard
Parch         | Number of parents/children aboard
Ticket        | Ticket number
Fare          | Fare
Cabin         | Cabin
Embarked      | Port of embarkation

Training data:
Total number of passengers - 891
Survival status distribution - 342/891=38.4% survived

```{r}
length(train$PassengerId)
table(train$Survived)
```
# Exploratory data analysis
The goal of this part of the analysis is to understand the data

## Possible predictors  
The following have strong impacts on Survival status:
* Pclass - higher class cabin passengers are more likely to survive
* Gender - female is much favored
* Age - kids (age < 10) are more possible to survive 59% for <10, >62 low chance 20%
* SibSp - more sibling/spouse, less chance to survive? >2.5 only 7/46 15%
* Parch - doesn't seem to have an impact
* People who travel alone - less chance to survive

The following don't seem to have impact on Survival status:
* Fare and Pclass are correlated, so no need to consider Fare
* Embarked
* Cabin - cabin number
* Ticket - ticket number

```{r}
table(train$Pclass, train$Survived)
ggplot(train,aes(x=Pclass,fill=factor(Survived)))+
  geom_histogram(stat="count")
```
```{r}
table(train$Sex, train$Survived)
ggplot(train,aes(x=Sex,fill=factor(Survived)))+
  geom_histogram(stat="count")
```
```{r}
ggplot(train,aes(x=Age,fill=factor(Survived)))+
  geom_histogram(binwidth=2)
  
train$kid <- ifelse (train$Age<=10, "Yes", "No")
table(train$kid, train$Survived)

train$old <- ifelse (train$Age>62, "Yes", "No")
table(train$old, train$Survived)
```
```{r}
ggplot(train,aes(x=Fare,fill=factor(Survived)))+
  geom_histogram(binwidth=10)
  
library(dplyr)
train %>% 
group_by(Pclass) %>% 
summarize(n=length(Fare), mean=mean(Fare), min=min(Fare), max=max(Fare))
```
```{r}
ggplot(train,aes(x=SibSp,fill=factor(Survived)))+
  geom_histogram(binwidth=1)
  
train$moreSibSp <- ifelse (train$SibSp>2.5, "Yes", "No")
table(train$moreSibSp, train$Survived)
```
```{r}
ggplot(train,aes(x=Parch,fill=factor(Survived)))+
  geom_histogram(binwidth=1)
```
```{r}
ggplot(train,aes(x=Embarked,fill=factor(Survived)))+
  geom_histogram(stat="count")
```

```{r}
train$single <- ifelse(train$SibSp==0 & train$Parch==0, "Yes", "No")
table(train$single, train$Survived)
```

